---
title: "Import / Export data to / from R"
author: "Partha Das"
date: "`r Sys.Date()`"
output: 
  html_document:
#    css: custom.css
    toc: true
    toc_float: true
#    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Intializing Libraries

```{r}
library(readr)
library(readxl)
library(openxlsx)
library(here)
```

# Read/ write delimited text files

A lots of data comes as a simple tabular text files. We shall learn how to to import those data into R using `readr` package of tidyverse family

## Reading single delimited files

Reading a comma separated file

```{r, message=FALSE}

netflix <- read_csv(here("cstep_tutorials/data", "netflix_titles_nov_2019.csv")); netflix

```

reading tab delimited files

```{r, message=FALSE}

netflix <- read_tsv(here("cstep_tutorials/data", "netflix_titles_nov_2019.txt")); netflix

```

Read any delimited files

```{r, message=FALSE}

netflix <- read_delim(here("cstep_tutorials/data", "netflix_sep_dollar.txt"), delim = '$'); netflix
```

### Exploring some arguments

* `col_names`
* `col_types`
* `na`
* `skip`
* `n_max`
* `guess_max`

### Parsing

## Writing a delimited file

Comma delimited using `write_csv()`

```{r}

write_csv(netflix, here('cstep_tutorials/out', 'netflix_comma_csv.csv'))
```

Tab delimited using `write_tsv()`

```{r}

write_tsv(netflix, here('cstep_tutorials/out', 'netflix_comma_csv.csv'))
```

Comma delimited using `write_delim()`

```{r}

write_delim(netflix, here('cstep_tutorials/out', 'netflix_comma_delim.txt'), delim= ',')
```

Tab delimited using `write_delim()`

```{r}

write_delim(netflix, here('cstep_tutorials/out', 'netflix_sep_tab.txt'), delim = '\t')
```

'\$' delimited using `write_delim()`

```{r}

write_delim(netflix, here('cstep_tutorials/out', 'netflix_sep_dollar.txt'), delim = '$')
```


# Read/ write Excel files

A lots of data comes as Excel spreadsheets simple tabular text files. We shall learn how to to import those data into R using `readr` package of tidyverse family

## Read Excel files with single sheet

```{r, message=FALSE}

netflix <- read_excel(here("cstep_tutorials/data", "netflix_titles_nov_2019.xlsx"), sheet = 1); netflix

netflix <- read_excel(here("cstep_tutorials/data", "netflix_titles_nov_2019.xlsx"), sheet = 'netflix_titles_nov_2019'); netflix

```

## Write Excel files

Single sheet

```{r, message=FALSE}

openxlsx::write.xlsx(netflix, here("cstep_tutorials/out", "netflix_titles_nov_2019.xlsx"), sheetName = 'netflix_1')

openxlsx::write.xlsx(netflix, here("cstep_tutorials/out", "netflix_titles_nov_2019.xlsx"), sheetName = 'netflix_2')

```

Look into the options

# Advanced import export

## Read multiple delimited files

## Read Excel files with multiple sheets

## Read multiple Excel files with multiple sheets

Multiple sheets

```{r, echo=FALSE, eval=FALSE}

# create a workbook

netflix_data = createWorkbook()

# create sheets
sheet1 = addWorksheet(netflix_data, 'netflix')
netflix_sheet = addWorksheet(netflix_data, 'netflix', gridLines = FALSE)

# write data into the sheets

# save the workbook
saveWorkbook(netflix_data, here("cstep_tutorials/out", "multi_sheet_example.xlsx"), overwrite = TRUE)

```

```{r, echo=FALSE}
list_of_datasets <- list(netflix, netflix)
write.xlsx(list_of_datasets, file = "writeXLSX2.xlsx", overwrite = TRUE)
```

